\section{ODD Protocol}
\label{app:odd}

This appendix documents EDEM in the standard ODD protocol for
agent-based models \citep{grimm2020odd}.

\subsection{Purpose and patterns}

EDEM's purpose is to characterise the conditions under which an
agent-based real-estate market reaches a stable equilibrium price,
and to identify the mechanisms that prevent equilibrium when they
do not. Patterns the model is designed to reproduce: stable
band-bounded equilibria; endogenous business cycles; persistent
shifts in the realised price relative to the textbook equilibrium
under altered patience or density; multiplicative price drift in the
absence of a balancer.

\subsection{Entities, state variables, and scales}

\paragraph{Entities.} \textbf{Buyers} and \textbf{Sellers}
(\texttt{mesa.Agent} subclasses) and \textbf{Homes} (per-cell
dataclass instances on a $32 \times 32$ toroidal grid).

\paragraph{State variables.}
\begin{itemize}[leftmargin=*,nosep]
\item Buyer: epsilon (estimation-error bound), delay, current bid,
      heading. EDEM variant additionally: lowest-bid-to-value ratio,
      yellow flag.
\item Seller: epsilon, patience timer, current ask price, dictionary
      of received bids. EDEM additionally: number of bids, total
      bids, best-bid-to-value, best-bid-to-true-value.
\item Home: market price, last-sale tick, last-sale price, fair value
      $v^{*}$, current value $v$.
\item Model: rolling window of last 25 sale prices (DE), epoch
      counter (EDEM), current cross-population epsilon (EDEM).
\end{itemize}

\paragraph{Spatial and temporal scales.} One spatial unit = one home;
one tick is the smallest temporal unit. Each tick triggers one round
of buyer movement and one round of seller bid-processing. DE balance
period is 100 ticks; EDEM epoch is 20 ticks.

\subsection{Process overview and scheduling}

Per tick, in order:
\begin{enumerate}[leftmargin=*,nosep]
\item Each agent steps once (in randomised order over both classes).
      A Buyer's step posts bids, then wiggles, then steps forward.
      A Seller's step processes patience and best-bid logic.
\item The model invokes the balancer:
      DE recomputes linear targets every 100 ticks;
      EDEM fires the per-epoch update every 20 ticks (using the
      cycle-counter trick of \cref{sec:implementation-clearing} to
      preserve buyer \texttt{is\_yellow} flags).
\item The data collector samples model-level reporters.
\end{enumerate}

\subsection{Design concepts}

\paragraph{Basic principles.} EDEM extends the dynamic
divergence-of-opinion premise of \citet{miller1977} to a
multi-period setting; agents act on noisy estimates of an unobservable
fair value.

\paragraph{Emergence.} The price level, agent population dynamics,
and macroscopic regimes (band-stable, business-cycle, persistent
shift, bubble, transitional) emerge from the local interaction of
agent estimation, bidding, and patience timers; none of these
phenomena are coded in directly.

\paragraph{Adaptation.} Sellers lower ask prices when their patience
timer elapses without high-enough bids; this is the only adaptive
behaviour. Buyers do not adapt: they place bids drawn from a fixed
distribution and accept by Cond.~2 (\cref{eq:cond2}).

\paragraph{Objectives.} Sellers maximise realised sale price
subject to patience; buyers minimise realised purchase price subject
to Cond.~2. Neither has an explicit utility function; objectives
are implicit in their action rules.

\paragraph{Learning.} None in the baseline model
(\cref{sec:ext-adaptive} sketches an adaptive extension).

\paragraph{Prediction.} None.

\paragraph{Sensing.} Buyers sense whether a seller is on their patch;
sellers sense the bids in their own bid table. Neither senses the
global market price directly.

\paragraph{Interaction.} Direct: a buyer that lands on a seller's
patch posts a bid via a mirrored-dictionary update on both sides.
Indirect: the average ask price (DE) and the rolling sale-price
average (DE) influence newly-spawned agents' starting prices.

\paragraph{Stochasticity.} Used for: agent placement at setup;
buyer heading initialisation and per-tick wiggle; epsilon draws
per agent; per-bid epsilon-error draws; victim selection in the
balancer; Bernoulli draws for fractional $C_{b}$ swaps in EDEM.
All draws come from a single seeded \texttt{numpy.random.Generator}
owned by the model.

\paragraph{Collectives.} None.

\paragraph{Observation.} The data collector records, per tick:
market price (DE: rolling-25 mean; EDEM: average home value),
agent counts, the equilibrium price implied by current schedules,
the rolling-window fill count, and (EDEM) the current epsilon.
Output is stacked into Parquet datasets, one per seed, for
downstream analysis.

\subsection{Initialisation}

DE: spawn \texttt{equi\_qnty} sellers at random unique cells with
patience $\sim \mathcal{U}[0, \texttt{init\_patience})$ and ask price
drawn from $\mathcal{U}[-\bar{\sigma}, +\bar{\sigma}]$ around
$p^{*}$; spawn \texttt{equi\_qnty} buyers at random cells.
EDEM: 20 sellers and 20 buyers, all homes initialised with
$v_{0}(h) = v^{*}(h) = 100$.

\subsection{Input data}

None. EDEM is a closed model; all dynamics are endogenous. The shock
hooks in Run~5 are scheduled in the experiment script, not loaded
from external data.

\subsection{Submodels}

\paragraph{Estimation function.} \cref{eq:estimation} with
$\varepsilon \sim \mathcal{U}[-\sigma_{i}, +\sigma_{i}]$ and
$\sigma_{i} \sim \mathcal{U}[0, \bar{\sigma}]$.

\paragraph{Bid acceptance.} \cref{eq:cond2}.

\paragraph{Market price.} DE: \cref{eq:price-rolling}, the rolling
mean of the last $W = 25$ sale prices. EDEM: \cref{eq:value-update},
the per-epoch multiplicative update.

\paragraph{Balancer.} DE: linear-target restoration every $T_{B}$
ticks. EDEM: \cref{eq:Qs-update,eq:Qd-update} in finite-population
form, with fractional $C_{b}$ realised as integer + Bernoulli swaps
and a population floor of one agent per side.

\paragraph{Agent spawn / exit.} On a sale, both counterparties leave
the market and a fresh pair is spawned by the balancer with patience
drawn from $\mathcal{U}[50, \texttt{init\_patience})$ for sellers
and immediate readiness for buyers.
